Geographically aware estimates of remotely sensed water properties for Chesapeake Bay

### 2022-09-19 manuscript/manuscript.tex
typo
multi-multi-panel prediction maps

Tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure S1). Out of sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1). Despite excellent overall performance, our temperature and salinity models had poor performance at the extremes. They were unable to recover very low values ($<$ 0 C) and very high values ($>$ 30 C, $>$ 23 salinity, Figure 1). These areas of poor performance can be explained by their relative rarity in the observational dataset (Table S2).

@@

Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance from each tributary mouth (Figure 3). RF model predictions were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 4). The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 2).

\begin{figure}[ht!]
    \begin{center}
          \includegraphics[width=0.4\textwidth,keepaspectratio]{figures/_seasonality}
          \caption{Seasonal Random Forest salinity prediction results.}
    \end{center}    
\end{figure}

@@

Another potential area of future research is resolving water properties at depths beyond the surface. A key challenge in this area is lack of observational data at multiple locations. Depth arrayed measurements are typically only associated with buoy deployments of which there are few in the Bay. Perhaps these can be supplemented by data from physics-based models such as CBOFS. The results of such as an effort can be regarded more so as an emulator than a standalone model.

>>>>

Tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure 1). Out of sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1). Despite excellent overall performance, our temperature and salinity models had poor performance at the extremes. They were unable to recover very low values ($<$ 0 C) or very high values ($>$ 30 C, $>$ 23 salinity, Figure 1). One explanation for these areas of poor performance are their relative rarity in the observational dataset (Table S2).

@@

Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance from each tributary mouth (Figure 3). A realistic spatial trend can also be seen in the RF temperature predictions whereby maximum water temperatures were located in the upper tributaries while minimum water temperatures were highest at the Bay mouth (Figure 3).

RF model predictions for salinity were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 4). The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 3).

\begin{figure}[ht!]
    \begin{center}
          \includegraphics[width=0.8\textwidth,keepaspectratio]{figures/_seasonality}
          \caption{Seasonal Random Forest prediction results for temperature and salinity.}
    \end{center}    
\end{figure}

@@

Another potential area of future research is resolving water properties at depths beyond the surface. A key challenge in this area is lack of observational data at multiple locations. Depth arrayed measurements are typically only associated with buoy deployments of which there are few in the Bay. Perhaps these can be supplemented by data from physics-based models such as CBOFS. The results of such an effort can be regarded more so as an emulator than a standalone model.
